Supplementary Material for Food-web complexity alters the fitness landscape of an insect herbivore

### 2018-05-03 supp_mat_complexity_selection.Rmd
created template for 2D fitness landscape plots.

Load dataset. This dataset is a subset that only includes larva that survive to pupation or turned into parasitoids. We omitted any data where larva mortality was unclear. This enabled us to focus on the selection pressures imposed by insect parasitoids rather than other sources.
```{r Preliminaries, message=FALSE, warning=FALSE, include=FALSE}
## LOAD REQUIRED LIBRARIES & SET OPTIONS ----
library(tidyverse)
library(cowplot) # pretty default ggplots
library(broom) # for tidying multiple linear models
library(mgcv) # generalized additive models
library(gsg) # calculating selection gradients
library(visreg)

@@

```{r FUNCTIONS}
gamm.model <- function(formula, data) {
  gamm4(formula, data, 
        random = ~(1|Genotype/Plant_Position/Gall_Number),
        family = binomial(link = logit))
}

@@

# experimenting with boot.case gave bootstrapped distributions that weren't centered near the model mean estimate.
# also, using 'refit.smooth = T' also generated some very skewed distributions
# the recipe below provided the most sensible estimates for my data
gradient.calc <- function(mod, phenotype, covariates = NULL) {
  gam.gradients(mod, phenotype, covariates, parallel = 'multicore', ncpus = 32, 
                refit.smooth = F, 
                n.boot = 1000,
                se.method = 'boot.para',
                standardized = T)
} 
```

@@

## Gall Height
gam_red.control_size <- gam(gall_survival ~
                             s(c.Gall_Height_mm) + 
                             c.gall_individuals +
                             c.Density_per_100_shoots,
                           data = control_df, family = binomial(link = "logit"))
summary(gam_red.control)

gam.plot(gam_red.control)
gam.plot(gamm_red.control$gam)

@@

gam_red.control_indiv <- gam(gall_survival_mean ~
                              c.Gall_Height_mm_mean + 
                              s(c.gall_individuals, k=9) +
                              c.Density_per_100_shoots,
                             data = control_df_gall.level, family = binomial(link = "logit")) # we didn't include weights because this would be correlated with the number of individuals per gall
summary(gam_red.control_indiv)

gam.plot(gam_red.control_indiv)
gam.plot(gamm_red.control$gam)

## Gall density
control_df_plant.level <- control_df_gall.level %>%
  group_by(Treatment.focus, Plant_Position) %>%
  summarise(gall_survival_mean = mean(gall_survival_mean),
            gall_sample_size = n(),
            gall_individuals_mean = mean(gall_individuals),                 
            Gall_Height_mm_mean = mean(Gall_Height_mm_mean),
            Density_per_100_shoots = mean(Density_per_100_shoots)) %>%                # note that there is no variance, it is the actual value
  ungroup() %>%
  mutate(c.Gall_Height_mm_mean = Gall_Height_mm_mean - mean(Gall_Height_mm_mean),
         c.gall_individuals_mean = gall_individuals_mean - mean(gall_individuals_mean),
         c.Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots)) # why is this giving zero values?

gam_red.control_density <- gam(gall_survival_mean ~
                                c.Gall_Height_mm_mean + 
                                c.gall_individuals_mean +
                                s(c.Density_per_100_shoots, k=3),   # do I adjust k to be consistent with mixed-effects model?
                             data = control_df_plant.level, family = binomial(link = "logit"), weights = gall_sample_size) # weighting appears to alter the results...
summary(gam_red.control_density)

gam.plot(gam_red.control_density)
gam.plot(gamm_red.control$gam)

@@

## Gall Height
gam_red.treatment_size <- gam(gall_survival ~
                         s(c.Gall_Height_mm) + 
                         c.gall_individuals +
                         c.Density_per_100_shoots,
                       data = treatment_df, family = binomial(link = "logit"))
summary(gam_red.treatment)

gam.plot(gam_red.treatment)
gam.plot(gamm_red.treatment$gam)

## Gall individuals
treatment_df_gall.level <- treatment_df %>%
  group_by(Treatment.focus, Plant_Position, Gall_Number) %>%
  summarise(gall_survival_mean = mean(gall_survival),
            gall_individuals = mean(gall_individuals),                 # note that there is no variance, it is the actual value
            Gall_Height_mm_mean = mean(Gall_Height_mm),
            Density_per_100_shoots = mean(Density_per_100_shoots)) %>% # remember that this is pseudo-replicated at the gall-level
  ungroup() %>%
  mutate(c.Gall_Height_mm_mean = Gall_Height_mm_mean - mean(Gall_Height_mm_mean),
         c.gall_individuals = gall_individuals - mean(gall_individuals),
         c.Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots)) # why is this giving zero values?

gam_red.treatment_indiv <- gam(gall_survival_mean ~
                              c.Gall_Height_mm_mean + 
                              s(c.gall_individuals, k=9) +
                              c.Density_per_100_shoots,
                             data = treatment_df_gall.level, family = binomial(link = "logit"))
summary(gam_red.treatment_indiv)

gam.plot(gam_red.treatment_indiv)
gam.plot(gamm_red.treatment$gam)

## Gall density
treatment_df_plant.level <- treatment_df_gall.level %>%
  group_by(Treatment.focus, Plant_Position) %>%
  summarise(gall_survival_mean = mean(gall_survival_mean),
            gall_sample_size = n(),
            gall_individuals_mean = mean(gall_individuals),                 
            Gall_Height_mm_mean = mean(Gall_Height_mm_mean),
            Density_per_100_shoots = mean(Density_per_100_shoots)) %>%                # note that there is no variance, it is the actual value
  ungroup() %>%
  mutate(c.Gall_Height_mm_mean = Gall_Height_mm_mean - mean(Gall_Height_mm_mean),
         c.gall_individuals_mean = gall_individuals_mean - mean(gall_individuals_mean),
         c.Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots)) # why is this giving zero values?

gam_red.treatment_density <- gam(gall_survival_mean ~
                                c.Gall_Height_mm_mean + 
                                c.gall_individuals_mean +
                                s(c.Density_per_100_shoots),
                             data = treatment_df_plant.level, family = binomial(link = "logit"), weights = gall_sample_size) # weighting appears to alter the results...
summary(gam_red.treatment_density)

gam.plot(gam_red.treatment_density)
gam.plot(gamm_red.treatment$gam)

@@

## GALL SIZE ----
sg.gam.size_control <- gradient.calc(mod = gam_red.control, 
                                 phenotype = "c.Gall_Height_mm", 
                                 covariates = c("c.gall_individuals","c.Density_per_100_shoots"))
sg.gam.size_treatment <- gradient.calc(mod = gam_red.treatment, 
                                   phenotype = "c.Gall_Height_mm", 
                                   covariates = c("c.gall_individuals","c.Density_per_100_shoots"))

## GALL INDIVIDUALS ----
sg.gam.individuals_control <- gradient.calc(mod = gam_red.control, 
                                        phenotype = "c.gall_individuals", 
                                        covariates = c("c.Gall_Height_mm","c.Density_per_100_shoots"))
sg.gam.individuals_treatment <- gradient.calc(mod = gam_red.treatment, 
                                          phenotype = "c.gall_individuals", 
                                          covariates = c("c.Gall_Height_mm","c.Density_per_100_shoots"))

## GALL DENSITY ----
sg.gam.density_control <- gradient.calc(mod = gam_red.control, 
                                    phenotype = "c.Density_per_100_shoots", 
                                    covariates = c("c.gall_individuals","c.Gall_Height_mm"))
sg.gam.density_treatment <- gradient.calc(mod = gam_red.treatment, 
                                      phenotype = "c.Density_per_100_shoots", 
                                      covariates = c("c.gall_individuals","c.Gall_Height_mm"))

@@

>>>>

Load dataset. This dataset is a subset that only includes larva that survive to pupation or turned into parasitoids. We omitted any data where larva mortality was unclear. This enabled us to focus on the selection pressures imposed by insect parasitoids rather than other sources.
```{r Preliminaries, message=FALSE, warning=FALSE, include=FALSE}
## LOAD REQUIRED LIBRARIES & SET OPTIONS ----
library(tidyverse)
library(cowplot) # pretty default ggplots
library(broom) # for tidying multiple linear models
library(mgcv) # generalized additive models
library(gamm4) # generalized additive mixed-models
library(gsg) # calculating selection gradients
library(visreg)
library(viridis) # for color palette

@@

```{r FUNCTIONS}
gamm.model <- function(formula, data) {
  gamm4(formula, data, 
        random = ~(1|Plant_Position/Gall_Number), # Genotype/
        family = binomial(link = logit))
}

@@

# experimenting with boot.case gave bootstrapped distributions that weren't centered near the model mean estimate.
# also, using 'refit.smooth = T' also generated some very skewed distributions
# the recipe below provided the most sensible estimates for my data
gradient.calc <- function(mod, phenotype, covariates = NULL) {
  gam.gradients(mod, phenotype, covariates, parallel = 'multicore', ncpus = 32, 
                refit.smooth = F, 
                n.boot = 1000,
                se.method = 'boot.para',
                standardized = T)
}

insect_FL <- function(mod, phenotype, covariates = NULL) {
  fitness.landscape(mod, phenotype, covariates, parallel = 'multicore', ncpus = 32,
                    PI.method = 'boot.para',
                    refit.smooth = F,
                    PI.interval = c(0.025,0.975),
                    plt.density = 100)
}

get_insect_FL.df <- function(FL){
  data.frame(mean_fitness = FL$Wbar, FL$points, t(FL$WbarPI)) %>%
    mutate(lower_2.5 = X2.5., upper_97.5 = X97.5.)
}

ggplot_uniFL <- function(FL.df) {
  ggplot(FL.df, aes(x = Var1, y = mean_fitness)) +
    geom_line() + 
    geom_ribbon(aes(ymin = lower_2.5, ymax = upper_97.5), alpha = 0.5)
}
```

@@

## Gall Height
gam_red.control_size <- gam(gall_survival ~
                             s(c.Gall_Height_mm) + 
                             s(c.gall_individuals, k=9) +
                             s(c.Density_per_100_shoots),
                           data = control_df, family = binomial(link = "logit"))
summary(gam_red.control_size)

gam.plot(gam_red.control_size, pages=1)
gam.plot(gamm_red.control$gam, pages=1)

@@

gam_red.control_indiv <- gam(gall_survival_mean ~
                              s(c.Gall_Height_mm_mean) + 
                              s(c.gall_individuals, k=9) +
                              s(c.Density_per_100_shoots),
                             data = control_df_gall.level, family = binomial(link = "logit")) # we didn't include weights because this would be correlated with the number of individuals per gall
summary(gam_red.control_indiv)

gam.plot(gam_red.control_indiv, pages=1)
gam.plot(gamm_red.control$gam, pages=1)

## Gall density
control_df_plant.level <- control_df_gall.level %>%
  group_by(Treatment.focus, Plant_Position) %>%
  summarise(gall_survival_mean = mean(gall_survival_mean),
            gall_sample_size = n(),
            gall_individuals_mean = mean(gall_individuals),                 
            Gall_Height_mm_mean = mean(Gall_Height_mm_mean),
            Density_per_100_shoots = mean(Density_per_100_shoots)) %>%    # note that there is no variance, it is the actual value
  ungroup() %>%
  mutate(c.Gall_Height_mm_mean = Gall_Height_mm_mean - mean(Gall_Height_mm_mean),
         c.gall_individuals_mean = gall_individuals_mean - mean(gall_individuals_mean),
         c.Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots))

#test.control_df_plant.level <- control_df %>%
#  group_by(Treatment.focus, Plant_Position) %>%
#  summarise(gall_survival_mean = mean(gall_survival),
#            gall_sample_size = n(),
#            gall_individuals_mean = mean(gall_individuals),                 
#            Gall_Height_mm_mean = mean(Gall_Height_mm),
#            Density_per_100_shoots = mean(Density_per_100_shoots)) %>%                # note that there is no variance, it is the actual value
#  ungroup() %>%
#  mutate(c.Gall_Height_mm_mean = Gall_Height_mm_mean - mean(Gall_Height_mm_mean),
#         c.gall_individuals_mean = gall_individuals_mean - mean(gall_individuals_mean),
#         c.Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots))

gam_red.control_density <- gam(gall_survival_mean ~
                                s(c.Gall_Height_mm_mean) + 
                                s(c.gall_individuals_mean, k=9) +
                                s(c.Density_per_100_shoots, k=3),  # adjusting k value to avoid overfitting of Density   
                             data = control_df_plant.level,
                             family = binomial(link = "logit"), 
                             weights = gall_sample_size) 
summary(gam_red.control_density)

gam.plot(gam_red.control_density, pages=1)
gam.plot(gamm_red.control$gam, pages=1)

@@

## Gall Height
gam_red.treatment_size <- gam(gall_survival ~
                         s(c.Gall_Height_mm) + 
                         s(c.gall_individuals, k=9) +
                         s(c.Density_per_100_shoots),
                       data = treatment_df, family = binomial(link = "logit"))
summary(gam_red.treatment_size)

gam.plot(gam_red.treatment_size, pages=1)
gam.plot(gamm_red.treatment$gam, pages=1)

## Gall individuals
treatment_df_gall.level <- treatment_df %>%
  group_by(Treatment.focus, Plant_Position, Gall_Number) %>%
  summarise(gall_survival_mean = mean(gall_survival),
            gall_individuals = mean(gall_individuals),              # note that there is no variance, it is the actual value
            Gall_Height_mm_mean = mean(Gall_Height_mm),
            Density_per_100_shoots = mean(Density_per_100_shoots)) %>% # remember this is pseudo-replicated at the gall-level
  ungroup() %>%
  mutate(c.Gall_Height_mm_mean = Gall_Height_mm_mean - mean(Gall_Height_mm_mean),
         c.gall_individuals = gall_individuals - mean(gall_individuals),
         c.Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots)) # why is this giving zero values?

gam_red.treatment_indiv <- gam(gall_survival_mean ~
                              s(c.Gall_Height_mm_mean) + 
                              s(c.gall_individuals, k=9) +
                              s(c.Density_per_100_shoots),
                             data = treatment_df_gall.level, family = binomial(link = "logit"))
summary(gam_red.treatment_indiv)

gam.plot(gam_red.treatment_indiv, pages=1)
gam.plot(gamm_red.treatment$gam, pages=1)

## Gall density
treatment_df_plant.level <- treatment_df_gall.level %>%
  group_by(Treatment.focus, Plant_Position) %>%
  summarise(gall_survival_mean = mean(gall_survival_mean),
            gall_sample_size = n(),
            gall_individuals_mean = mean(gall_individuals),                 
            Gall_Height_mm_mean = mean(Gall_Height_mm_mean),
            Density_per_100_shoots = mean(Density_per_100_shoots)) %>%  # note that there is no variance, it is the actual value
  ungroup() %>%
  mutate(c.Gall_Height_mm_mean = Gall_Height_mm_mean - mean(Gall_Height_mm_mean),
         c.gall_individuals_mean = gall_individuals_mean - mean(gall_individuals_mean),
         c.Density_per_100_shoots = Density_per_100_shoots - mean(Density_per_100_shoots))

gam_red.treatment_density <- gam(gall_survival_mean ~
                                s(c.Gall_Height_mm_mean) + 
                                s(c.gall_individuals_mean, k=9) +
                                s(c.Density_per_100_shoots),
                             data = treatment_df_plant.level, family = binomial(link = "logit"), 
                             weights = gall_sample_size) 
summary(gam_red.treatment_density)

gam.plot(gam_red.treatment_density, pages=1)
gam.plot(gamm_red.treatment$gam, pages=1)

@@

## GALL SIZE ----
sg.gam.size_control <- gradient.calc(mod = gam_red.control_size, 
                                 phenotype = "c.Gall_Height_mm", 
                                 covariates = c("c.gall_individuals","c.Density_per_100_shoots"))
sg.gam.size_treatment <- gradient.calc(mod = gam_red.treatment_size, 
                                   phenotype = "c.Gall_Height_mm", 
                                   covariates = c("c.gall_individuals","c.Density_per_100_shoots"))

## GALL INDIVIDUALS ----
sg.gam.individuals_control <- gradient.calc(mod = gam_red.control_indiv, 
                                        phenotype = "c.gall_individuals", 
                                        covariates = c("c.Gall_Height_mm_mean","c.Density_per_100_shoots"))
sg.gam.individuals_treatment <- gradient.calc(mod = gam_red.treatment_indiv, 
                                          phenotype = "c.gall_individuals", 
                                          covariates = c("c.Gall_Height_mm_mean","c.Density_per_100_shoots"))

## GALL DENSITY ----
sg.gam.density_control <- gradient.calc(mod = gam_red.control_density, 
                                    phenotype = "c.Density_per_100_shoots", 
                                    covariates = c("c.gall_individuals_mean","c.Gall_Height_mm_mean"))
sg.gam.density_treatment <- gradient.calc(mod = gam_red.treatment_density, 
                                      phenotype = "c.Density_per_100_shoots", 
                                      covariates = c("c.gall_individuals_mean","c.Gall_Height_mm_mean"))

@@

```{r Fitness Landscapes}
## GALL SIZE ----
fl.size_control <- insect_FL(mod = gamm_red.control$gam, 
                                 phenotype = "c.Gall_Height_mm", 
                                 covariates = c("c.gall_individuals","c.Density_per_100_shoots"))
fl.size_treatment <- insect_FL(mod = gamm_red.treatment$gam, 
                                   phenotype = "c.Gall_Height_mm", 
                                   covariates = c("c.gall_individuals","c.Density_per_100_shoots"))

## GALL INDIVIDUALS ----
fl.individuals_control <- insect_FL(mod = gamm_red.control$gam, 
                                        phenotype = "c.gall_individuals", 
                                        covariates = c("c.Gall_Height_mm","c.Density_per_100_shoots"))
fl.individuals_treatment <- insect_FL(mod = gamm_red.treatment$gam, 
                                          phenotype = "c.gall_individuals", 
                                          covariates = c("c.Gall_Height_mm","c.Density_per_100_shoots"))

## GALL DENSITY ----
fl.density_control <- insect_FL(mod = gamm_red.control$gam, 
                                    phenotype = "c.Density_per_100_shoots", 
                                    covariates = c("c.gall_individuals","c.Gall_Height_mm"))
fl.density_treatment <- insect_FL(mod = gamm_red.treatment$gam, 
                                      phenotype = "c.Density_per_100_shoots", 
                                      covariates = c("c.gall_individuals","c.Gall_Height_mm"))

get_insect_FL.df(fl.size_control) %>%
  ggplot_uniFL(.)

get_insect_FL.df(fl.size_treatment) %>%
  ggplot_uniFL(.)

get_insect_FL.df(fl.individuals_control) %>%
  ggplot_uniFL(.)

get_insect_FL.df(fl.individuals_treatment) %>%
  ggplot_uniFL(.)

get_insect_FL.df(fl.density_control) %>%
  ggplot_uniFL(.)

get_insect_FL.df(fl.density_treatment) %>%
  ggplot_uniFL(.)

## Landscape
fl.individuals.x.density_control <- insect_FL(mod = gamm_red.control$gam, 
                                              phenotype = c("c.Density_per_100_shoots","c.gall_individuals"), 
                                              covariates = c("c.Gall_Height_mm"))

get_insect_FL.df(fl.individuals.x.density_control) %>%
  ggplot(., aes(x=Var1, y=Var2, fill=mean_fitness)) +
  geom_tile()

fl.individuals.x.density_treatment <- insect_FL(mod = gamm_red.treatment$gam, 
                                                phenotype = c("c.Density_per_100_shoots","c.gall_individuals"), 
                                                covariates = c("c.Gall_Height_mm"))
get_insect_FL.df(fl.individuals.x.density_treatment) %>%
  ggplot(., aes(x=Var1, y=Var2, fill=mean_fitness)) +
  geom_tile()

test <- bind_rows(mutate(get_insect_FL.df(fl.individuals.x.density_treatment), Treatment.focus="Simple"),
          mutate(get_insect_FL.df(fl.individuals.x.density_control), Treatment.focus="Complex"))

ggplot(test, aes(x=Var1, y=Var2, fill=mean_fitness)) +
  geom_raster() + 
  facet_wrap(~Treatment.focus) +
  scale_fill_viridis()
```
